Parsing Postfix log files

### 2008-10-28 doc/logparser-other-parsers-review.tex
Minor changes.
First pass through corrections for sections 1-3.

Ten other parsers, starting with the only previously published research in
this area, have been reviewed as part of the background research for this
project.  None of the parsers reviewed perform the kind of advanced parsing
and log correlation required for this project; all bar one are intended to
perform a specific parsing and reporting task, rather than to be a generic
parser, extracting data and leaving generation of reports from the data to
other programs.  Some parsers save the data extracted to a data store, but
the majority discard all data once they have finished running, making
historical analysis impossible.  The other parsers reviewed all produce a
report of varying complexity and detail, whereas the program described here
does not produce a report at all; that responsibility is deferred to a
separate program, to be developed later.  The parsing algorithm and program
developed for this project are designed to enable much more detailed log
analysis by providing a stable platform for subsequent programs to develop
upon.

@@

There only appears to be one prior paper published about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL or Berkeley DB database.  The decision to
support both \CSV{} and Berkeley DB appears to have been a serious
limitation: \CSV{} is a very simple format where a record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with \CSV{} files include the difficulty in escaping separators,
providing multiple values for a field (e.g.\ multiple recipients) and
adding new fields.  There is no mechanism to document the fields or the
separator, unlike \SQL{} databases where every database includes a schema
naming the fields and the type of data they store (integer, text,
timestamp, etc.).

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \SQL{} databases that store arbitrary tuples; in \LMA{}'s main
table the key is an integer referred to by secondary tables, and the value
used is a \CSV{} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date and by-\IP{} tables use the
sender/recipient/date/\IP{} as the key, and a \CSV{} list of integers
referring to the main table is the value.  This effectively re-implements
\SQL{} foreign keys, but without the functionality offered by even the most
basic of \SQL{} databases (joins, ordering, searches, etc.).  It also
requires custom code to search on some combination of the above, though the
authors of \LMA{} did provide some queries: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  Berkeley DB appears to be the least useful
of the three output formats: it does not provide the functionality of a
basic \SQL{} database, not can it be used with standard Unix text
processing tools.

The data stored is limited to time and date, hostname and \IP{} address of
client, \IP{} of server, sender and recipient addresses, \SMTP{} code, and
size (for accepted mails only).  Handling of multiple recipients, \SMTP{}
codes or remote servers\footnote{A single mail may be sent to multiple
remote servers if it was addressed to recipients in different domains, or
Postfix needs to try multiple servers for one or more recipients.} is not
explained in the paper; experimental observation shows that multiple
records are added when there are multiple recipients (sadly the records are
not associated or linked in any way), and presumably the same approach is
taken when there are multiple destination servers.

@@

Parsing in \LMA{} requires major changes to the code to parse new log lines
or extract additional data.  It does not appear to deal with any of the
complications discussed in \sectionref{complications} (initial
complications) and \sectionref{additional complications} (additional
complications).  It does not differentiate between different types of
rejections, so it is not suitable for the purposes of this project; the
data about which restriction caused the rejection is discarded, whereas an
important goal of this project is to retain that data to aid optimisation
and evaluation of Postfix restrictions.  \LMA{} fails to parse Postfix log
files generated on Solaris hosts because the fields added automatically
differ from those added on Linux hosts; log files from Solaris hosts (and
possibly other operating systems) thus require preprocessing before parsing
by \LMA{}.  Once the preprocessing has been performed on the
\numberOFlogFILES{} test log files \LMA{} parses the log files without
complaint, though it does produce 32 entries in its output file for every
rejection in the input log file; it also missed some 40\% of delivered
mail.  Once these glaring deficiencies were discovered the author did not
spend more time checking the results.

\LMA{} does provide some simple reports: IP-STORY, FROM-STORY, DAILY-EMAIL
and DAILY-REJECT\@.  IP-STORY shows the \CSV{} lines with the specified
client \IP{} address; FROM-STORY shows the \CSV{} lines with the specified
sender address; DAILY-EMAIL shows the \CSV{} lines for the specified day.
DAILY-REJECT initially failed with an error message from the Perl
interpreter; after making some corrections to the code it worked, and
produced the \CSV{} lines for the specified day where the \SMTP{} code
signifies a rejection.  All of these reports are trivially simple to
produce from the \CSV{} file using the standard Unix tool \texttt{awk},
e.g.\ the most complicated, DAILY-REJECT, is merely:

@@

    \item It outputs an unnecessary line containing only a single space
        after each matching line, for compatibility with the output of
        DAILY-REJECT\@.

    \item Some records have an \SMTP{} code of \textit{0}, though this is
        not a valid \SMTP{} code; presumably this is because of \LMA{} not
        parsing a line properly.  \LMA{} considers this code to be a
        rejection, so the command above must, too.

    \item In a tiny proportion of records the \SMTP{} code is
        \textit{deferred\/} (again, not a valid code, and most likely
        because of mis-parsing by \LMA{}); it is treated as a rejection by
        \LMA{}, so the equivalent \texttt{awk} command must do so, too.

@@

There are two differences in the output from DAILY-REJECT and the
\texttt{awk} command:

@@

    \item \LMA{} does not escape or replace the separators used in the
        \CSV{} records when they are present in sender or recipient
        addresses, leading to some mangled records; these records are not
        present in the output from \texttt{awk}.

    \item Where there are multiple records in the \CSV{} file, DAILY-REJECT
        produces only one output line, whereas the \texttt{awk} command
        above produces all matching lines.

@@

In summary, \LMA{} appears to be a proof of concept, written to demonstrate
the point of their paper (that having this information in an accessible
fashion is useful to systems administrators), rather than a program
designed to be extensible and useful in a production environment.

@@

It is important to compare and contrast newly developed programs,
algorithms and parsers against those already available, to accurately judge
what, if any, improvements are delivered by the newcomers.  There are not
that many previously developed Postfix log parsers --- indeed it was quite
difficult to find ten parsers to review for this project --- and the
functionality offered ranges from quite basic to much more mature,
depending on the needs of the creator.  The programs are not reviewed from
an independent viewpoint; the objective is to compare and contrast each
program with \parsername{}.  There are some important differences between
\parsername{} and the parsers reviewed here:

    \item Only \parsername{} makes it possible to parse new log lines
        without modifying the core parsing algorithm; see \sectionref{why
        separate rules, actions and framework?} for a discussion of why the
        separation of the algorithm into rules, actions and framework is
        beneficial.

    \item \parsername{} does not provide any reporting based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        log files and extracting data --- further processing of that data
        is a separate concern, deferred to another program.  Using an
        \SQL{} database simplifies the process of generating such reports
        (discussed in \sectionref{database as API}); some simple examples
        are given in \sectionref{motivation}.

    \item Most of the parsers silently ignore log lines they cannot handle,
        whereas \parsername{} complains loudly about every single line it
        fails to parse.  The exception is AWStats, which outputs the
        percentage of input lines it was unable to parse, but does not
        output the lines themselves.

@@

Pflogsumm produces a report designed for troubleshooting, rather than for
in-depth analysis.  It does not support saving any data or extracting more
data than is required for producing the report.  Both the parsing and
reporting are difficult to extend as it is a specialised tool, unlike the
easily extensible design of \parsername{}.  It does not correlate log lines
by queueid or \pid{}, and does not need to deal with the complications
encountered during this project.  Pflogsumm produces a useful report, and
successfully dealt with the \numberOFlogFILES{} log files tested
with,\footnote{The results it reported were not verified, but it did not
report any errors, and has a very good reputation amongst Postfix users.}
but is not a suitable base for this project.

@@

    \item Total number of mails accepted, delivered and rejected.  Total
        bytes of mail accepted and delivered.

@@

    \item Per-hour averages and per-day summaries of the number of mails
        received, delivered, deferred, bounced and rejected.

    \item For received mail: per-domain totals for mails sent, deferred,
        average delay, maximum delay and bytes delivered.  For received
        mail: per-domain totals for mails received and bytes received.

@@

\url{http://jimsun.linxnet.com/postfix_contrib.html} \newline (Last checked
2008/04/09.)

@@

Sawmill is a general purpose commercial product that parses 687 log file
formats (correct as of 2007/12/04) and produces reports.  Its data
extraction facilities are too limited to save sufficient data for the
purposes of this project,\footnote{The data extracted by Sawmill is
described later in this review, and the data extracted by \PLP{} is
described in \sectionref{connections table} and \sectionref{results
table}.} though it does extract three different sets of data for Postfix
(one is beta as of 2007/12/04), but they do not appear to be interlinked.
No attempt is made to correlate log lines or deal with the difficulties
documented in \sectionref{complications} \sectionref{additional
complications}.\footnote{If any attempt is made there is no reference to it
in the documentation available on the website.} The source code is
available in an obfuscated form only (presumably for a fee), and the
product is quite expensive, as it requires a license per report to be
generated; in contrast \parsername{} is free and the code is freely
available.  The web interface allows creation of dynamic reports based on
any field, but because of the associated cost the author has not
experimented with it.  Presumably, as the company charges per report the
user wishes to generate, the data store (if there is one) is deliberately
inaccessible and undocumented to prevent the user bypassing the program and
generating their own reports.

\url{http://www.thesawmill.co.uk/formats/postfix.html} \newline Fields
extracted: from, to, server, UID, relay, status, number of recipients,
origin hostname, origin \IP{} and virus.  The fields \texttt{server},
\texttt{uid} and \texttt{virus} are not explained in their documentation:
\texttt{server} is probably the server the mail is delivered to, and
\texttt{uid} might be the uid of the user submitting mail locally.  Postfix
does not perform any form of virus checking (though it has many options for
cooperating with an external virus scanner), so the \texttt{virus} field is
a mystery.

\url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline Fields
extracted: from, to, RBL, client hostname and client \IP{}\@.

\url{http://www.thesawmill.co.uk/formats/beta_postfix.html} \newline Fields
extracted: from, to, client hostname, client \IP{}, relay hostname, relay
\IP{}, status, response code, RBL and message id.

Even if the three data sets were linked together Sawmill would extract less
data than \parsername{}, and it does not appear to extract data about
rejections except when the rejection is caused by an \DNSBL{} check.

When tested with the \numberOFlogFILES{} test log files it performed
adequately, though the rate it processed log files at did slow down
noticeably as it progressed.  Sawmill supports reading compressed log files
but it exhibits a dramatic slow down when doing so: it took six hours to
parse the first half of the log files; twelve hours to parse the next
third; after twenty four hours parsing the remaining sixth it crashed due
to lack of disk space.  On the second parsing attempt the log files were
uncompressed beforehand and parsing took eight hours.

@@

(Last checked 2007/12/04.)

@@

Splunk aims to index all an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Saved
searches\footnote{The author was unable to save searches, though that may
have been due to limitations in the free version.} can be run periodically
and the results emailed to a recipient or sent to a shell script, that
presumably can publish the results as required (though possibly without the
graphs and detailed statistics); \parsername{} provides the database and
leaves it to the user to utilise it in any way, whenever they want, with no
restrictions on usage of the data.  The interface is optimised for
interactive rather than automated queries and it does not appear to be
possible to write independent tools to utilise the Splunk database; there
are additional reports available on \url{http://www.splunkbase.com/}, but
the author was unable to find any: every category was empty, even those
that the interface claimed had reports available.  Many types of reports
are bundled with the software, though most are variations of a bar or pie
chart, except bubble and heatmap graphs.  It is easy to drill down through
the graphs to extract a portion of the data (e.g.\ select the hour with the
largest number of events, then select a particular host, and finally a
specific address), though it is not possible to search on partial words.

The web interface is quite attractive and simple to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing it took roughly 30 minutes to add a single log file to
be indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all log files in a
particular directory, but after three hours, numerous futile attempts, and
reading all available documentation the author admitted defeat.  Using the
\CLI{} was more successful: invoking the command \newline \tab{}
\texttt{splunk find logs }\textit{log-directory\/}\newline added 40 of the
\numberOFlogFILES{} log files to the queue for indexing.  Repeated attempts
enqueued the same 40 log files, without explaining why the others
were excluded.\footnote{The log files appear to have only been indexed
once; presumably Splunk keeps track of the files it has indexed and
discards requests to index files for a second time.  This may or may not be
a useful feature for \parsername{}.} There did not appear to be an option
to order the log files to ensure they would be processed in the order they
were created.  Subsequently the author was successful in adding a single
file at a time using the \CLI{}; a simple loop to add all desired log files
was sufficient to index all files.  Splunk will periodically check all
those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.  All
searches performed using the indexed data returned reasonable results.

@@

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available that allows
indexing of up to 5GB of data per day.  The cheapest licensed version costs
\$5000 (plus \$1000 support), and still limits the volume of data indexed
per day to 500MB\@.

When parsing Postfix log files Splunk extracts some standard fields: to and
from addresses, HELO hostname, date, host the log files were collected
from, and protocol.  It parses the standard syslog fields at the beginning
of the line, and extracts any \texttt{key=<value>} pairs occurring after
the standard syslog prologue; these pairs are the fields listed above.
Searches can be based on the extracted fields, and all the text in the line
is also available for searching.  \parsername{} extracts noticeably more
data, though it does not make the full text of the line available, and the
full power of \SQL{} is available when searching, allowing the user to
search on arbitrarily complicated conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields; most importantly it makes
no attempt to correlate log lines by queueid or \pid{}, or to handle any of
the myriad complications discussed in this document.  Some additional
Postfix reports are supposedly available at
\url{http://www.splunkbase.com/}, but the author was unable to find them
(or any other reports).

\url{http://www.splunk.com/} \newline (Last checked 2008/04/29.)

@@

Isoqlog produces a report listing the number of mails sent by each unique
sender address, and separately the total number of bytes transferred; both
reports are produced for daily, monthly and annual time spans, but only for
the domains listed in its configuration file (making it impossible to
produce reports for every sender domain).  It appears to ignore all log
lines except for those for the current day, though it does maintain a
record of data previously extracted, which the newly extracted data is
merged into (no information is provided on the format of the data store).
The data extracted appears to be limited to the number of mails sent by
each sender, unlike the copious amounts of data extracted by \parsername{}.
It does not utilise rejection log lines in any way, so is unsuitable for
the purposes of this project.  Its parsing is completely inextensible,
indeed is almost incomprehensible, relying on \texttt{scanf(3)}, fixed
offsets and low level string manipulation; it is the opposite end of the
spectrum to \parsernames{} parsing.  It does not handle any of the
complications discussed in this document, does not gather the breadth of
data required for this project, and ignores the majority of log lines
produced by Postfix.  It ignores all log lines earlier than today, so
testing with the test log files is pointless.

\url{http://www.enderunix.org/isoqlog/} \newline (Last checked 2007/08/13.)

@@

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code restricts
it to supporting the \LCD{}.  The data it will extract from an \MTA{} log
file is limited in comparison to \parsername{}: \newline \tab{} time2,
email, email\_r, host, host\_r, method, url, code and bytesd.\newline There
does not seem to be an explanation of any of those fields in the
documentation (\parsername{} provides copious documentation).  AWStats has
no special purpose code to deal with the intricacies of Postfix log files,
in fact it operates by coercing Postfix log files into Apache\footnote{The
Apache web server is the most popular HTTP server in use for the past 10
years; more information is available at \url{http://httpd.apache.org/}.}
format log files, for analysis by AWStats' HTTP log file parser.  The
converting parser only deals with a small portion of the log lines
generated by Postfix, silently skipping those it cannot deal with, and does
not distinguish between different types of rejection; it would be a lot of
work to extend it to handle new log lines.  Although though it does
correlate log lines by queueid, it does not deal with any of the other
complications described in this document.  AWStats supports saving data but
the format of the saved data is not documented (as far as the author could
tell).  It also supports reading compressed log files, but that
functionality was not tested.

When tested with the \numberOFlogFILES{} test log files AWStats' report
says it parsed 9,240,075 (88.70\%) log lines out of 10,416,129, with
1,176,050 (11.29\%) corrupt lines; however there are actually
\numberOFlogLINES{} lines in the \numberOFlogFILES{} log files, so AWStats
only parsed 17.15\% of the input lines, ignoring the remaining 82.85\%.
The graphs it produces give an overview of mails received for the last
calendar month, showing:

@@

\url{http://anteater.drzoom.ch/} \newline (Last checked 2007/08/13.)

>>>>

Ten other parsers, starting with the only previously published research in
this area that the author is aware of, have been reviewed as part of the
background research for this project.  None of the parsers reviewed perform
the kind of advanced parsing and correlation required for this project; all
except one (XXX IS THIS ACCURATE\@?) are intended to perform a specific
parsing and reporting task, whereas \PLP{} is designed to be a generic
parser, extracting data and leaving generation of reports from the data to
other programs.  Some parsers save the data extracted to a data store, but
the majority discard all data once they have finished running, making
historical analysis impossible.  The parsers reviewed all produce a report
of varying complexity and detail, whereas this project's parser does not
produce a report at all; that responsibility is deferred to a separate
program, to be developed later.  The parser developed for this project is
designed to enable much more detailed log analysis by providing a stable
platform for subsequent programs to develop upon.

It is important to compare and contrast newly developed programs,
algorithms, and parsers against those already available, to accurately
judge what improvements, if any, are delivered by the newcomers.  There are
not that many previously developed Postfix log parsers --- indeed it was
quite difficult to find ten parsers to review for this project --- and the
functionality offered ranges from quite basic to much more mature,
depending on the needs of the creator.  The programs are not reviewed from
an independent viewpoint; the objective is to compare and contrast each
program with \parsername{}.  There are some important differences between
\parsername{} and the parsers reviewed here:

    \item Only \parsername{} makes it possible to parse new log lines
        without extensive and intrusive modifications to the parser;
        \parsername{}'s architecture is described in \sectionref{why
        separate rules, actions and framework?}.

    \item \parsername{} does not provide any reports based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        log files and extracting data; further processing of that data is a
        separate concern, deferred to another program.  Using an \SQL{}
        database simplifies the process of generating such reports
        (discussed in \sectionref{database as API}); some simple examples
        are given in \sectionref{motivation}.

    \item Most of the parsers silently ignore log lines they cannot handle,
        whereas \parsername{} complains loudly about every single log line
        it fails to parse.  The exception is AWStats, which outputs the
        percentage of input lines it was unable to parse, but does not
        output the lines themselves.

@@

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL or Berkeley DB database.  The decision to
support both \CSV{} and Berkeley DB appears to have been a serious
limitation: \CSV{} is a very simple format where a record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with \CSV{} files include the need to escape separators in the
data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  There is no mechanism to document the
fields or the separator, unlike \SQL{} databases where every database
includes a schema naming the fields and the type of data they store
(integer, text, timestamp, etc.).

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \SQL{} databases that store arbitrary tuples.  In \LMA{}'s main
table the key is an integer referred to by secondary tables, and the value
is a \CSV{} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date, and by-\IP{} tables use the
sender/recipient/date/\IP{} as the key, and the value is a \CSV{} list of integers
referring to the main table.  This effectively re-implements
\SQL{} foreign keys, but without the functionality offered by even the most
basic of \SQL{} databases (joins, ordering, searches, etc.).  It also
requires custom code to search on some combination of the above, though the
authors of \LMA{} did provide some queries: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  Berkeley DB appears to be the least useful
of the three output formats: it does not provide the functionality of a
basic \SQL{} database, not can it be used with standard Unix text
processing tools.

\LMA{} stores the following date: time and date, hostname and \IP{} address
of client, \IP{} of server, sender and recipient addresses, \SMTP{} code,
and size (for accepted mails only).  It omits XXX XXX\@.  Handling of
multiple recipients, \SMTP{} codes or remote servers\footnote{A single mail
may be sent to multiple remote servers if it was addressed to recipients in
different domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained in the paper; experimental observation shows
that multiple records are added when there are multiple recipients (sadly
the records are not associated or linked in any way), and presumably the
same approach is taken when there are multiple destination servers.

@@

\LMA{} requires major changes to the parser code to parse new log lines or
extract additional data.  The code is structured as a long series of blocks
that each handle all lines matching a single \regex{}, so parsing new log
lines requires modifying an existing \regex{} or carefully inserting a new
block in the correct place; extracting extra data will require modifying
multiple blocks, \regexes{}, or both.

\LMA{} does not appear to deal with any of the complications discussed in
\sectionref{complications}.  It does not differentiate between different
types of rejections, so it is not suitable for the purposes of this
project; the data about which restriction caused the rejection is
discarded, whereas the main goal of this project is to retain that data to
aid optimisation and evaluation of Postfix restrictions.  \LMA{} fails to
parse Postfix log files generated on Solaris hosts because the fields
automatically prepended to each log line differ from those added on Linux
hosts; log files from Solaris hosts (and possibly other operating systems)
thus require preprocessing before parsing by \LMA{}.  Once the
preprocessing has been performed on the \numberOFlogFILES{} test log files
\LMA{} parses the log files without complaint, although it produced 32
entries in its output file for every rejection in the input log file; it
also missed some 40\% of delivered mail.  Once these glaring deficiencies
were discovered the author did not waste any more time checking the
results.

\LMA{} does provide some simple reports: IP-STORY, FROM-STORY, DAILY-EMAIL
and DAILY-REJECT\@.  The first three extract \CSV{} lines for the specified
client \IP{} address, sender address, or date respectively.  DAILY-REJECT
initially failed with an error message from the Perl interpreter; after
making some corrections to the code it worked, and extracted the \CSV{}
lines for the specified day where the \SMTP{} code signifies a rejection.
All of these reports are trivially simple to produce from the \CSV{} file
using the standard Unix tool \texttt{awk}; the most complicated,
DAILY-REJECT, is merely:

@@

    \item It outputs an unnecessary line containing only a single space
        after each matching line, to accurately replicate the output of
        DAILY-REJECT\@.

    \item Some records have an \SMTP{} code of \textit{0}, though this is
        not a valid \SMTP{} code; presumably this is because \LMA{} did not
        parse a log line properly.  DAILY-REJECT considers this code to be
        a rejection, so the command above must too.

    \item In a small fraction of records the \SMTP{} code is
        \textit{deferred\/} (again, not a valid \SMTP{} code, and most
        likely because of mis-parsing by \LMA{}); it is treated as a
        rejection by DAILY-REJECT, so the equivalent \texttt{awk} command
        must do so, too.

@@

There are two differences between the output from DAILY-REJECT and the
\texttt{awk} command:

@@

    \item \LMA{} does not escape or replace the separators used in the
        \CSV{} records when they are present in sender or recipient
        addresses, leading to some mangled records; these records are not
        present in the output from \texttt{awk}.  XXX WHY\@?  GIVE AN
        EXAMPLE RECORD\@.

    \item Where there are multiple records in the \CSV{} file, DAILY-REJECT
        extracts only one record, whereas the \texttt{awk} command above
        extracts all matching records.

@@

In summary, \LMA{} appears to be a proof of concept, written to demonstrate
the point of their paper (that having this information in an accessible
fashion is useful to systems administrators), rather than a program
designed to be useful in a production environment; no consideration was
given to extensibility when writing it.

@@

@@

\noindent{}\url{http://jimsun.linxnet.com/postfix_contrib.html} \newline{}
(Last checked 2008/04/09.)

Pflogsumm produces a report designed for troubleshooting rather than for
in-depth analysis.  It does not support saving any data or extracting more
data than is required for producing the report.  Both the parsing and
reporting are difficult to extend as it is a specialised tool, unlike the
easily extensible design of \parsername{}.  It does not correlate log lines
by queueid or \pid{}, and does not need to deal with the complications
encountered during this project.  Pflogsumm produces a useful report, and
successfully parsed the \numberOFlogFILES{} log files tested
with,\footnote{The results it reported were not verified in detail, but it
did not report any errors, and has a very good reputation amongst Postfix
users.} but it is not a suitable base for this project.

@@

    \item Total number and bytes of mails accepted, delivered and rejected.

@@

    \item Per-hour averages and per-day summaries of the number of mails
        received, delivered, deferred, bounced, and rejected.

    \item For received mail: per-domain totals for mails sent, deferred,
        average delay, maximum delay, and bytes delivered.  For received
        mail: per-domain totals for mails received and bytes received.

@@

@@

\noindent{}\url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{}
\url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline{}
\url{http://www.thesawmill.co.uk/formats/beta_postfix.html} \newline{}
(Last checked 2007/12/04.)

Sawmill is a general purpose commercial product that parses 687 log file
formats (correct as of 2007/12/04) and produces reports.  Its data
extraction facilities are too limited to save sufficient data for the
purposes of this project,\footnote{The data extracted by Sawmill is
described later in this review, and the data extracted by \PLP{} is
described in \sectionref{connections table} and \sectionref{results
table}.} though it does extract three different sets of data for Postfix
(one is beta as of 2007/12/04), but they do not appear to be interlinked.
The documentation does not suggest that any attempt is made to correlate
log lines or deal with the difficulties documented in
\sectionref{complications}.  The source code is available in an obfuscated
form only (presumably for a fee), and the product is quite expensive,
requiring a license per report to be generated; in contrast \parsername{}
is free and the code is freely available.  The web interface allows
creation of dynamic reports based on any field, but because of the
associated cost the author has not experimented with it.  Presumably, as
the company charges per report the user wishes to generate, the data store
(if there is one) is deliberately inaccessible and undocumented to prevent
the user bypassing the program and generating their own reports.

\url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{} Fields
extracted: from, to, server, UID, relay, status, number of recipients,
origin hostname, origin \IP{} and virus.  The fields \texttt{server},
\texttt{uid} and \texttt{virus} are not explained in the documentation:
\texttt{server} is probably the server the mail is delivered to, and
\texttt{uid} might be the uid of the user submitting mail locally.  Postfix
does not perform any form of virus checking (though it has many options for
cooperating with an external virus scanner), so the \texttt{virus} field is
a mystery.

\url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline{} Fields
extracted: from, to, RBL, client hostname and client \IP{}\@.

\url{http://www.thesawmill.co.uk/formats/beta_postfix.html} \newline{} Fields
extracted: from, to, client hostname, client \IP{}, relay hostname, relay
\IP{}, status, response code, RBL and message id.

Even if the three data sets were linked together Sawmill would extract less
data than \parsername{}, and it does not appear to extract data about
rejections except when the rejection is caused by a \DNSBL{} check.

When tested with the \numberOFlogFILES{} test log files it performed
adequately, though the rate it processed log files at did slow down
noticeably as it progressed.  Sawmill supports reading compressed log files
but it exhibits a dramatic slow down when doing so: it took six hours to
parse the first half of the log files, and twelve hours to parse the next
third; after twenty four hours parsing the remaining sixth it crashed due
to lack of disk space.  On the second parsing attempt the log files were
uncompressed beforehand and parsing took eight hours.

@@

XXX ADD A SUMMARY\@.

@@

\noindent{}\url{http://www.splunk.com/} \newline{}
(Last checked 2008/04/29.)

Splunk aims to index all an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Saved searches can
be run periodically and the results emailed to a recipient or sent to an
external program for further processing (though maybe without the graphs
and detailed statistics); the author was unable to save searches, though
that may have been due to limitations in the free version.  The database is
not available for use by external programs, whereas \parsername{} provides
the database and leaves it to the user to utilise it without limit or
restriction.  The interface is optimised for interactive rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database; there are additional
reports available on \url{http://www.splunkbase.com/}, but the author was
unable to find any: every category was empty, even those that the interface
claimed had reports available.  Many types of reports are bundled with the
software, though most are variations of a bar or pie chart, except bubble
and heatmap graphs.  It is easy to drill down through the graphs to extract
a portion of the data (e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific address),
though it is not possible to search on partial words.

The web interface is quite attractive and simple to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing it took roughly 30 minutes to add a single log file to
be indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all the log files in
a particular directory, but after three hours, numerous futile attempts,
and reading all the available documentation the author admitted defeat.
Using the \CLI{} was more successful: invoking the command \newline{}
\tab{} \texttt{splunk find logs }\textit{log-directory\/}\newline{} added
40 of the \numberOFlogFILES{} log files to the queue for indexing.
Repeated attempts enqueued the same 40 log files, without explaining why
the others were excluded.\footnote{The log files appear to have only been
indexed once; presumably Splunk keeps track of the files it has indexed and
discards requests to index files for a second time.  This may or may not be
a useful feature for \parsername{}.} There did not appear to be an option
to order the log files to ensure they would be processed in the order they
were created.  Subsequently the author was successful in adding a single
file at a time using the \CLI{}; a simple loop to add all desired log files
was sufficient to index all files.  Splunk will periodically check all
those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.  All
searches performed using the indexed data returned reasonable results.

@@

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available that allows
indexing of up to 5GB of data per day.  The cheapest licensed version costs
\$5000 (plus \$1000 support), and still limits the volume of data indexed
per day to 500MB\@.  Typical log file sizes for a small scale mail server
are given in \sectionref{parser efficiency}.

When parsing Postfix log files Splunk extracts some standard fields: to and
from addresses, HELO hostname, time and date, the host the log files were
generated on, and protocol (\SMTP{} or \ESMTP{}).  It parses the standard
syslog fields at the beginning of the log line, and extracts any
\texttt{key=<value>} pairs occurring after the standard syslog prologue;
these pairs are the fields listed above.  Searches can be based on the
extracted fields or the full text of the log line.  \parsername{} extracts
noticeably more data, though it does not make the full text of the line
available, and the full power of \SQL{} is available when searching,
allowing the user to search on arbitrarily complicated conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields from a log line; most
importantly it makes no attempt to correlate log lines by queueid or
\pid{}, or to handle any of the myriad complications discussed in
\sectionref{complications}.  Some additional Postfix reports are supposedly
available at \url{http://www.splunkbase.com/}, but the author was unable to
find Postfix reports, or additional reports for any other log file types.

@@

\noindent{}\url{http://www.enderunix.org/isoqlog/} \newline{}
(Last checked 2007/08/13.)

Isoqlog's report lists daily, monthly, and annual totals of mails sent by
each unique sender and bytes transferred (XXX EXPLAIN BYTES TRANSFERRED\@:
TO/FROM, PER-SENDER\@?).  It only produces these reports for the domains
listed in its configuration file, making it impossible to produce reports
for every sender domain.  It ignores all log lines except those for today,
so it is impossible to analyse historical logs, and testing with the
\numberOFlogFILES{} test log files is pointless.  It does maintain a record
of data previously extracted, which the newly extracted data is merged into
(no information is provided on the format of the data store).  The data
extracted is limited to the number of mails sent by each sender, unlike the
copious amounts of data extracted by \parsername{}.  It does not utilise
rejection log lines in any way, so is unsuitable for the purposes of this
project.  Its parsing is completely inextensible, indeed is almost
incomprehensible, relying on \texttt{scanf(3)}, unexplained fixed offsets,
and low level string manipulation; it is the opposite end of the spectrum
to \parsernames{} parsing.  It does not handle any of the complications
discussed in \sectionref{complications}, does not gather the breadth of
data required for this project, and ignores the majority of log lines
produced by Postfix.

XXX FINISH THIS\@.

@@

\noindent{}\url{http://awstats.sourceforge.net/} \newline{}
\url{http://awstats.sourceforge.net/awstats.mail.html} \newline{}
\url{http://awstats.sourceforge.net/docs/awstats_faq.html#MAIL} \newline{}
(Last checked 2007/10/13.)

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code limits its
functionality.  The data it will extract from an \MTA{} log file is limited
in comparison to \parsername{}: \newline{} \tab{} time2, email, email\_r,
host, host\_r, method, url, code, and bytesd.\newline{} XXX WHAT FIELDS ARE
MISSING\@?  There does not seem to be an explanation of any of those fields
in the documentation (\parsername{} provides copious documentation).
AWStats coerces Postfix log files into Apache\footnote{The Apache web
server is the most popular HTTP server in use over the past 10 years; more
information is available at \url{http://httpd.apache.org/}.} format log
files, for analysis by AWStats' HTTP log file parser.  The converting
parser only deals with a small portion of the log lines generated by
Postfix, silently skipping those it cannot deal with, and does not
distinguish between different types of rejection; it would be a lot of work
to extend it to handle new log lines.  Although it does correlate log lines
by queueid (not by pid), it does not deal with any of the other
complications described in \sectionref{complications}.  AWStats supports
saving data but the format of the saved data is not documented, as far as
the author could tell.  It also supports reading compressed log files, but
that functionality was not tested.

When tested with the \numberOFlogFILES{} test log files AWStats' reported
that it parsed 9,240,075 (88.70\%) of 10,416,129 log lines, skipping
1,176,050 (11.29\%) corrupt log lines; however there are
\numberOFlogLINES{} lines in the \numberOFlogFILES{} log files, so AWStats
parsed only 17.15\% of the input lines, ignoring the remaining 82.85\%.

The graphs it produces give an overview of mails received for the last
calendar month, showing:

@@
